%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Copyright (C) 2010-2012  Pietro Abate <pietro.abate@pps.jussieu.fr>   %
%                           Ralf Treinen <ralf.treinen@pps.jussieu.fr>   %
%                           Unversité Paris-Diderot                      %
%                                                                        %
%  This documentation is free software: you can redistribute it and/or   %
%  modify it under the terms of the GNU General Public License as        %
%  published by the Free Software Foundation, either version 3 of the    %
%  License, or (at your option) any later version.                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Output}
\label{sec:output}
The output of \debcheck{} is in the YAML format, see
Section~\ref{sec:tricks-python} for how to parse the output.

Without any particular options, \debcheck{} just reports some
statistics:
\begin{example}
\begin{verbatim}
% dose-debcheck rep1
background-packages: 0
foreground-packages: 4
total-packages: 4
broken-packages: 1
\end{verbatim}
\end{example}

With the options \texttt{--failures} and \texttt{--successes}, \debcheck{}
reports findings of the requested kind for all packages in the foreground.
These options may be used alone or in combination. In any case, the status
field tells whether the package is found to be installable (value \texttt{ok})
or non-installable (value \texttt{broken}).

\begin{example}
\begin{verbatim}
% dose-debcheck --failures --successes rep1
report:
 -
  package: a
  version: 1
  architecture: amd64
  source: a (= 1)
  status: broken
  
 -
  package: a
  version: 2
  architecture: amd64
  source: a (= 2)
  status: ok
  
 -
  package: b
  version: 1
  architecture: amd64
  source: b (= 1)
  status: ok
  
 -
  package: c
  version: 3
  architecture: amd64
  source: c (= 3)
  status: ok
  
 
background-packages: 0
foreground-packages: 4
total-packages: 4
broken-packages: 1
\end{verbatim}
\end{example}

With an additional \texttt{--explain} option, an explanation is given
with each finding. 

\subsection{Understanding Explanations of installability}

Explanations of installability are simply giving by listing an
installation (in the sense of Section~\ref{sec:installability})
containing the package in question.

\begin{example}
\begin{verbatim}
% dose-debcheck --explain --successes rep1
report:
 -
  package: a
  version: 2
  architecture: amd64
  source: a (= 2)
  status: ok
  installationset:
   -
    package: c
    version: 3
    architecture: amd64
   -
    package: a
    version: 2
    architecture: amd64
 -
  package: b
  version: 1
  architecture: amd64
  source: b (= 1)
  status: ok
  installationset:
   -
    package: b
    version: 1
    architecture: amd64
\end{verbatim}
\end{example}

\subsection{Understanding Explanations of Non-installability}

Installability of a package is much easier to explain than
non-installabilty. The reason for this is that in the former case we
just have to give one installation that our tool has found, while in
the latter case we have to explain why \emph{all} possible attempts to
install the package must fail. The first consequence of this
observation is that the explanation in case non-installability may
consist of several components.

\begin{example}
  Consider the following repository consisting of only two packages:
\begin{verbatim}
Package: a
Version: 1
Depends: b | c

Package: c
Version: 3
Conflicts: a
\end{verbatim}
To explain why package (\texttt{a},1) is not installable we have to
say why all possible alternative ways to satisfy its dependency must
fail:
\begin{itemize}
\item there is no package \texttt{b} in the repository
\item the only version of package \texttt{c} in the repository is in
  conflict with package (\texttt{a},1)
\end{itemize}
\end{example}

There may be several ways to satisfy dependencies due to alternatives
in the dependencies in packages. Alternatives may occur in dependencies
in different forms:
\begin{itemize}
\item explicitly, like in \texttt{Depends: b|c},
\item through dependency on a package that exists in several versions,
\item through dependency on a virtual package which is provided by several
  (versions of) real packages.
\end{itemize}
There is one component in the explanation for every possible way to
choose among these alternatives in the dependencies.

There are only two possible reasons why an attempt to satisfy dependencies
may fail:
\begin{enumerate}
\item dependency on a package that is missing from the repository,
\item dependency on a package that is in conflict with some other package
  we depend on (possibly through a chain of dependencies).
\end{enumerate}
Each component of the explanation is either a missing package, or a conflict. 

\subsubsection{Explanation in Case of a Missing Package}
A component of the explanation that corresponds to the case of a
missing consist of two stanzas:
\begin{itemize}
\item a \texttt{pkg} stanza that states the package that cannot satisfy
  one of its direct dependencies
\item a \texttt{paths} stanza containing the dependency chain that
  leads from the package we have found non-installable to the one that
  cannot satisfy its direct dependency.
\end{itemize}
\begin{example}
An explanation might look like this:
\begin{verbatim}
package: libgnuradio-dev
version: 3.2.2.dfsg-1
architecture: all
source: gnuradio (= 3.2.2.dfsg-1)
status: broken
reasons:
   -
    missing:
     pkg:
      package: libgruel0
      version: 3.2.2.dfsg-1+b1
      architecture: amd64
      unsat-dependency: libboost-thread1.40.0 (>= 1.40.0-1)
     paths:
      -
       depchain:
        -
         package: libgnuradio-dev
         version: 3.2.2.dfsg-1
         Architecture: all
         Depends: libgnuradio (= 3.2.2.dfsg-1)
        -
         package: libgnuradio
         ersion: 3.2.2.dfsg-1
         architecture: all
         depends: libgnuradio-core0
        -
         package: libgnuradio-core0
         version: 3.2.2.dfsg-1+b1
         architecture: amd64
         depends: libgruel0 (= 3.2.2.dfsg-1+b1)
\end{verbatim}
This tells us that \texttt{libgnuradio-dev} in version $3.2.2.dfsg-1$
is not installable, due to the fact that package \texttt{tlibgruel0}
in version $3.2.2.dfsg-1+b1$ has a dependency
\texttt{libboost-thread1.40.0 (>= 1.40.0-1)} that is not matched by
any package in the repository. The dependency chain tells why package
\texttt{libgnuradio-dev} in the given version might want to install
\texttt{libgruel0}.
\end{example}

The paths component gives all possible depchains from the root package
(\texttt{libgnuradio-dev} in the above example) to the one where a
direct dependency is not matched by any package (\texttt{libgruel0} in
the example). We do not include the last node in the dependency chain
to avoid a useless repetition. In general there may be more then one
path to reach a certain package from a given root package, in that
case \debcheck{} will unroll all of them.

\subsubsection{Explanation in Case of a Conflict}
The other possible cause of a problem is a conflict. In that case, the
explanation consists of a \texttt{conflict} stanza giving the two
packages that are in direct conflict with each other. Next, we have
two \texttt{depchain} stanzas that lead to the first, resp. the second
of these directly conflicting packages.
\begin{example}
\begin{verbatim}
package: a
  version: 1
  status: broken
  reasons:
   -
    conflict:
     pkg1:
      package: e
      version: 1
     pkg2:
      package: f
      version: 1
     depchain1:
      -
       depchain:
        -
         package: a
         version: 1
         depends: b
        -
         package: b
         version: 1
         depends: e
     depchain2:
      -
       depchain:
        -
         package: a
         version: 1
         depends: d
        -
         package: d
         version: 1
         depends: f
\end{verbatim}
The first part of the distcheck report is as before with details about the
broken package. Since this is a conflict, and all conflicts are binary, we give
the two packages involved in the conflict first. Packages f and e are in
conflict, but they are not direct dependency of package a . For this reason, we
output the two paths that from a lead to f or e. All dependency chains for each
conflict are together. Again, since there might be more than one way from a to
reach the conflicting packages, we can have more then one depchain. 
\end{example}
If a conflict occurs between two packages that are both reached
through non-trivial dependency chains then we sometimes speak of a
\emph{deep conflict}.



